Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-17 submit/fix.tex
new slew of simulations and results

\documentclass{aastex61}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{color}
\defcitealias{1999PASP..111...63F}{F99}
\defcitealias{2017ApJ...842...93M}{M17}
\linenumbers

@@

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
This approach is taken by \citet{2011A&A...529L...4C}, who find that after standardization based on Ca and Si features, remaining residual color
variation is consistent with Milky Way dust models.
We accommodate up to  three independent color parameters: the model is constructed such that 
two are assigned as extrinsic processes attributed to dust, and one is attributed to intrinsic supernova
properties.
The data used in this analysis are described in \S\ref{data:sec}.
In \S\ref{model:sec} we present a
first analysis with two extrinsic parameters and omitting the intrinsic parameter.  The properties of the two extrinsic parameters
are consistent with the expectations of parameterized Milky Way extinction models.
In \S\ref{model2:sec} we add the intrinsic parameter in the analysis, and find strong evidence for its influence on observed magnitudes.
In  \S\ref{discussion:sec}
we relate our results to known
signatures of SN~Ia diversity.
Conclusions are presented in \S\ref{conclusions:sec}.

@@

\section{Supernova Model I: Two Extrinsic Parameters}
\label{model:sec}

@@

Tests have revealed no evidence for the non-convergence of the MCMC run.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
(in this paragraph not to be confused with the synthetic $R$ magnitude)
\citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that all are densely sampled.
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig} -- \ref{global5:fig}.  In these tests there is no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
For four selected colors, a plot of the
model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
 versus observed colors and the histogram of their differences
are shown in Figure~\ref{residual:fig}. 
There is no evidence of catastrophically poor model predictions nor of extreme
outlying data.
We have rerun the analysis with a variety of initial conditions, including one with $\vec{\gamma}$ equal to zero, except for a small positive 
$\gamma^0_{\hat{U}}$ and small negative $\gamma^1_{\hat{U}}$;
the 68\% credible intervals
remain consistent.  Note that the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $k$.

@@

\subsection{Per-Supernova Parameter -- Covariance Matrix Degeneracy}
\label{degenerate:sec}
\color{blue}
As discussed in \S\ref{modelmodel:sec}, our two models have degeneracy between per-supernova parameters and the covariance matrix.
In the posterior, the degeneracy is broken by the determinant of the covariance matrix in the likelihood and by the priors.
Given that the priors we use are not exactly correct,
the parameters are susceptible to the misassignment, in either direction, between intrinsic color, $\Delta$, $k$, and $D$, and intrinsic dispersion $C_c$.
\S\ref{priorappendix:sec}

>>>>

\documentclass[trackchanges]{aastex61}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{color}
\defcitealias{1999PASP..111...63F}{F99}
\defcitealias{2017ApJ...842...93M}{M17}
\linenumbers

@@

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
This approach is taken by \citet{2011A&A...529L...4C}, who find that after standardization based on Ca and Si features, remaining residual color
variation is consistent with Milky Way dust models.
We accommodate up to  three independent color parameters: the model is constructed such that 
two are assigned as extrinsic processes attributed to dust, and one is attributed to intrinsic supernova
properties.
The data used in this analysis are described in \S\ref{data:sec}.
\replaced{
In \S\ref{model:sec} we present a
first analysis with two extrinsic parameters and omitting the intrinsic parameter.  The properties of the two extrinsic parameters
are consistent with the expectations of parameterized Milky Way extinction models.
In \S\ref{model2:sec} we add the intrinsic parameter in the analysis, and find strong evidence for its influence on observed magnitudes.
}{wefwe}
In  \S\ref{discussion:sec}
we relate our results to known
signatures of SN~Ia diversity.
Conclusions are presented in \S\ref{conclusions:sec}.

@@

\section{Model~I: Extrinsic Colors}
\label{modelI:sec}
We begin by considering a model for supernova magnitudes that have linear dependence on spectral parameters
and two extrinsic color parameters. This model is referred to as Model~I.  It is not the primary focus of this article, but its
results are of interest.  The purpose of this Section is to introduce some basic concepts of our models,
to show how the models can be expressed in terms of dust-extinction models, and to give Model~I results on dust. 
We refrain from presenting the full results of Model~I in deference to the more complex Model~II that is the subject of Section~\ref{modelII:sec},

\subsection{Model}
We assume 
that  peak intrinsic ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{Ca}$ and $EW_{Si}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{Si}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{Si}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
The intrinsic  magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}
\end{pmatrix},
\label{ewsiv:eqn}
\end{equation}
where $\vec{I}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) )$,.  This 
notationally distinguishes between global versus per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$, 
$\vec{\alpha}$, $\vec{\beta}$,  and $\vec{\eta}$,  are the intercept
(units of mag) and slopes (units of mag~\AA$^{-1}$)
of the linear relationships that
relate spectral-feature parameters with intrinsic magnitudes.

The apparent magnitudes are the sum of the intrinsic magnitudes plus additional terms that are linearly dependent on the
per-supernova
extrinsic-color parameters $g_0$ and $g_1$.  
Unlike the parameters associated with
the spectral measurements ($EW_{Ca}$, $EW_{Si}$ and $\lambda_{Si}$)  the latent,
dimensionless
parameters $g_0$ and $g_1$ are not directly associated
with observables but rather are inferred as part of the analysis.
The global parameter vectors $\vec{\gamma}^0$, $\vec{\gamma}^1$  are the slopes (units of mag) that relate 
the latent extrinsic-color parameters with apparent magnitudes.
The observables
$\hat{U}_o, {\hat{B}}_o, {\hat{V}}_o, {\hat{R}}_o, {\hat{I}}_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
have Gaussian measurement uncertainty with covariance $C$.
The
likelihood density for the described model
corresponds to the probability distribution function described by
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}

The model as written has degeneracies that need to be constrained in order for fits to converge.
The model does not specify the absolute magnitude nor the intrinsic color of a supernova that has no extrinsic color contribution;
in their place the zeropoints of the per-supernova magnitude and color parameters are set to the sample means
\begin{equation}
\langle \Delta \rangle=0,\ \langle g_0 \rangle=0,\ \langle g_1 \rangle=0.
\end{equation}
The model contains the product of latent parameters
$\vec{\gamma} g$, which leads to the degeneracy $\vec{\gamma} \rightarrow a\vec{\gamma}$, $g \rightarrow a^{-1} g$.
To aid in the convergence of $\vec{\gamma}$ and $g$ we impose a prior on the rms of $g$, though
our physical interpretations are ultimately independent of this scaling.
This prior does not specify the sign of $a$, which leaves a parity degeneracy.  As will be seen, the signal-to-noise in $\vec{\gamma}$ is sufficiently
high that our finite MCMC chains do not migrate between the degenerate solutions; to simplify the merging of multiple chains
we impose one of the degenerate solutions
\begin{equation}
\gamma^0_{\hat{U}} > 0,\ \gamma^1_{\hat{U}} < 0.
\end{equation}
As  in mixture models, the combinations $\vec{\gamma}^0 g_0$ and $\vec{\gamma}^1 g_1$ terms are degenerate under exchange of the 0 and 1 indices. For a cleaner presentation of our results, 
we break that degeneracy  by setting consistent initial conditions that succeed in keeping indices consistent over all analysis chains.

Another form of degeneracy is illustrated by the  transformations $\Delta \rightarrow \Delta  + \epsilon_\Delta$,
 $\vec{\gamma}^0 \rightarrow \vec{\gamma}^0  + \epsilon_{\gamma^0}$, $\vec{\gamma}^1 \rightarrow \vec{\gamma}^1 + \epsilon_{\gamma^1}$
 under the condition
$$
\epsilon_\Delta  +  \epsilon_{\gamma^0} g_0+  \epsilon_{\gamma^1} g_1=0,
$$
i.e.\ shifts in the global coefficients $\vec{\gamma}$ can be compensated by shifts in 
$\Delta$
for each supernova while maintaining 
$\langle \Delta \rangle=0$.
Another degeneracy is illustrated by $g_0 \rightarrow g_0 + (EW_{Ca}-\langle EW_{Ca}\rangle)\epsilon$,
$\vec{\alpha} \rightarrow \vec{\alpha} - \vec{\gamma} \epsilon$, $\vec{c} \rightarrow \vec{c} + \langle EW_{Ca}\rangle \epsilon$, i.e. the
linear external terms can leak into the linear spectral feature corrections.
These degeneracies are not broken within the model, but rather are constrained by the priors in which the
per-supernova parameters
$\Delta$, $EW$, $\lambda$, $g_0$, and $g_1$ are uncorrelated.

Except where stated otherwise above, we adopt a flat prior for all parameters.

\subsection{Results: $\gamma^0$ and $\gamma^1$}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Attenuation of supernova light due to dust is expected to be mostly captured by the first-order affects described by our linear model.
Here we  interpret our model $\gamma$'s in terms of
the dust-extinction model of \citetalias{1999PASP..111...63F}.

In the analysis of this and other sections of this article,
the posterior of the model parameters is evaluated using Hamiltonian Monte Carlo with a No-U-Turn
Sampler as implemented in
STAN \citep{stan}.  We run eight chains, each with 5000 iterations of which
half are used for warmup.
Tests have revealed no evidence for the non-convergence of the MCMC run.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
(in this paragraph not to be confused with the synthetic $R$ magnitude)
\citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that all are densely sampled.

All elements of the two $\vec{\gamma}$ parameter vectors 
describing intrinsic color
are significantly non-zero.
None of the 20000 links of 
our Monte Carlo chains for $\vec{\gamma}$ come close to the origin, we thus claim detection of the
influence of $g_0$ and $g_1$  on supernova magnitudes
with probability $1-5\times 10^{-5}$.
We thus find a strong signal for a linear 
extrinsic contribution to our model
\begin{equation}
A_X = \gamma^0_X g_0 +  \gamma^1_X g_1.
\end{equation}

A linear model with parameters $A_V$ and $E(B-V)$
\begin{equation}
A_X = a(X)  A_{V} + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes  
dust-extinction
models.  For example,   the wavelength-dependent model of  \citet{1989ApJ...345..245C} is linear,
though its linearity is lost when integrated over broad-band filters.
For the case of
$R^F=2.5$ and $A^F_V=0.1$ dust attenuating light from the SALT2
\citep{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at ${\hat{B}}$-band peak, the 
\citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F} model
can be approximated with
$\vec{a} = (0.96,   1.00,   1.00,   0.97,   0.77)$ and $\vec{b}=(  1.77,   0.98,   0.12,  -0.50,  -0.53)$.
(The $F$ superscript is used to distinguish parameters of the \citetalias{1999PASP..111...63F} model.)
Over the ranges
 $0\le A^F_V\le 1$ and $2 \le R^F \le 3.5$ 
and the wavelengths under consideration here,  the values of the elements of $a$ and $b$ vary by $<5$\%
for the
\citetalias{1999PASP..111...63F} model with
 the largest error in extinction from the linear approximation having amplitude $<0.008$ mag.

Our model $\gamma$ vectors can be written as a linear combination (described by the $2 \times 2$ matrix M) of the
dust-based $a(X)$ and $b(X)$ vectors plus residual vectors perpendicular to $\vec{a}$ and $\vec{b}$,
\begin{equation}
\begin{pmatrix}
\vec{\gamma}^0 \\
\vec{\gamma}^1
\end{pmatrix}=
M
\begin{pmatrix}
\vec{a} \\
\vec{b}
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_{\gamma^0} \\
\vec{\epsilon}_{\gamma^1}
\end{pmatrix}.
\label{trans_I:eqn}
\end{equation}
The $\gamma$'s are almost entirely captured by the first term. 
The Model~I posterior gives a quadratic fraction of $0.9993^{+0.0002}_{-0.0002}$ of $\vec{\gamma}^0$ projected onto the plane defined by
$a(X)$ and $b(X)$, the quadratic fraction of $\vec{\gamma}^1$ projected
onto the dust plane is
$0.959^{+0.013}_{-0.024}$.
This result indicates that
the allowed variations in ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are confined to almost identical
2-dimensional
planes within
the 5-dimensional magnitude space.
With no prior assumptions of dust extinction behavior or the distribution of $A_V$, the supernova data themselves exhibit
2-dimensional color variations that are closely aligned with the 2-dimensional color variations predicted by the \citetalias{1999PASP..111...63F} dust model.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{../output_fix1/plane0.pdf}
   \includegraphics[width=2.95in]{../output_fix1/plane1.pdf}
   \caption{
   Visualization of how supernova magnitudes can vary in our model and that of \citetalias{1999PASP..111...63F}.  While the models describe
   magnitudes in 5-bands ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ this visualization shows only ${\hat{U}}{\hat{V}}{\hat{I}}$.   The left and right plots show the same information from
   two different perspectives.
   Our model is shown in solid-line unit vectors in the directions of $\vec{\gamma}^0$ and $\vec{\gamma}^1$. The only possible magnitudes
   are a linear combination of these two vectors, and hence are confined to the plane that contains both.
   The  \citetalias{1999PASP..111...63F} model is shown in dashed-line unit vectors
   in the directions of  $\vec{a}$, $\vec{b}$; the only possible magnitudes are confined to the plane that contains both.
   The perspective on the right visually shows that all four vectors are nearly coplanar.  
   The combination $\vec{a}+\vec{b}/2.18$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\vec{\gamma}^0$.
   \label{plane:fig}}
\end{figure}

We
now turn to
the specific choice of $\vec{\gamma}$  returned by the fit within the context of the \citetalias{1999PASP..111...63F}  dust-extinction model. 
Our linear model would be satisfied by any two independent vectors  that span
the plane defined by the $\vec{\gamma}$ vectors.  While
Eq.~\ref{f99:eqn} expresses that model using basis vectors $\vec{a}$ and $\vec{b}$ and
their corresponding parameters 
$A_V$ and 
$E(B-V)$, the linear dust model 
alternatively
could be written in terms of any two other basis vectors  $\vec{a}+\kappa_1 \vec{b}$
and $\kappa_2 \vec{a} + \vec{b}$ as parameterized by $\kappa_1$ and $\kappa_2$.
In this basis, the model is then
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a(X)+\kappa_1 b(X) \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a(X) + b(X))],
\label{newdust:eqn}
\end{equation}
with the new parameters $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$  equal to  linear combinations of the original parameters.
Comparing this equation with  Eq.~\ref{trans_I:eqn}, we see that $\kappa_1=M[1,2]/M[1,1]$, $\kappa_2=M[2,1]/M[2,2]$.

Our posterior
includes a prior in which the per-supernova parameters $k_0$ and  $k_1$ associated with $\gamma$ are uncorrelated.
The parameters $\kappa_1$ and $\kappa_2$ are set by this prior.
If the external parameters of our model were indeed described by the \citetalias{1999PASP..111...63F}  dust model, then the prior corresponds
to    $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$ being uncorrelated in the supernova sample.
A trivial scenario that produces non-correlation is one in which all supernovae have the same parameter value for $-\kappa_1 A^F_V + E^F(B-V)$.
All color changes would then be confined to the $a(X)+\kappa_1 b(X)$ direction. This scenario corresponds to the \citetalias{1999PASP..111...63F}  dust model restricted to a constant $R^F=\kappa_1^{-1}$.   We use this interpretation to assign  $\langle R^F_{\mathit{eff}}\rangle = \kappa_1^{-1}$ as
the
ensemble mean
effective total-to-selective extinction of our sample.
With our posterior we
get $\kappa^{-1}_1=2.18^{+0.14}_{-0.16}$.
Figure~\ref{plane:fig} visualizes the near-perfect alignment of $\vec{\gamma}^0$ and  $\vec{a}+\vec{b}/2.18$.

Analysis of this model produces $\gamma$ vectors that can be
transformed to the Fitzpatrick dust-extinction model with a corresponding $2.02 <R< 2.32$ 68\%-ile credible interval.

Suppose supernova color are determined by more than two parameters.  For example, supernovae have some intrinsic color dispersion
in addition to being obscured by two-parameter dust.  A feature of this model is that the two color vectors $\gamma$ try to capture all these effects.

\section{Model~II: Extrinsic--Intrinsic Colors}
\label{modelII:sec}
Model~II introduces an intrinsic color dispersion described by a 1-dimensional Normal distribution.  The supernova population
is described by an eigenvector $\vec{\phi}$ and the color offset of each supernova is parameterized by
\begin{equation}
p \sim \mathcal{N}(0, \sigma_p).
\end{equation}
The model reads
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} + \phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  + \phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} + \phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} + \phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+ \phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} k_0 +\gamma^1_{{\hat{U}}} k_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} k_0 +\gamma^1_{{\hat{B}}} k_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} k_0+\gamma^1_{{\hat{V}}} k_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} k_0 + \gamma^1_{{\hat{R}}} k_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} k_0+\gamma^1_{{\hat{I}}} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Parameters& $X={\hat{U}}$ &${\hat{B}}$&${\hat{V}}$&${\hat{R}}$&${\hat{I}}$\\ \hline
$\alpha_X$
& $0.0048^{+0.0010}_{-0.0012}$
& $0.0019^{+0.0009}_{-0.0010}$
& $0.0018^{+0.0007}_{-0.0008}$
& $0.0017^{+0.0006}_{-0.0006}$
& $0.0027^{+0.0005}_{-0.0005}$
\\
${\alpha_X/\alpha_V-1}$
& $   1.7^{+   1.0}_{  -0.4}$
& $   0.1^{+   0.1}_{  -0.1}$
& $   0.0^{+   0.0}_{   0.0}$
& $  -0.0^{+   0.1}_{  -0.1}$
& $   0.5^{+   0.7}_{  -0.3}$
\\
$\beta_X$
& $ 0.040^{+ 0.003}_{-0.003}$
& $ 0.031^{+ 0.002}_{-0.002}$
& $ 0.033^{+ 0.002}_{-0.002}$
& $ 0.027^{+ 0.002}_{-0.002}$
& $ 0.027^{+ 0.002}_{-0.002}$
\\
${\beta_X/\beta_V-1}$
& $  0.23^{+  0.04}_{ -0.04}$
& $ -0.04^{+  0.02}_{ -0.02}$
& $  0.00^{+  0.00}_{  0.00}$
& $ -0.16^{+  0.01}_{ -0.01}$
& $ -0.18^{+  0.03}_{ -0.03}$
\\
$\eta_X$
& $0.0020^{+0.0010}_{-0.0009}$
& $0.0022^{+0.0009}_{-0.0008}$
& $0.0031^{+0.0008}_{-0.0008}$
& $0.0031^{+0.0007}_{-0.0007}$
& $0.0025^{+0.0007}_{-0.0007}$
\\
${\eta_X/\eta_V-1}$
& $ -0.35^{+  0.17}_{ -0.22}$
& $ -0.29^{+  0.10}_{ -0.14}$
& $  0.00^{+  0.00}_{  0.00}$
& $  0.01^{+  0.06}_{ -0.04}$
& $ -0.19^{+  0.10}_{ -0.09}$
\\
$\gamma^0_X$
& $ 71.32^{+  3.40}_{ -3.24}$
& $ 58.31^{+  3.03}_{ -2.93}$
& $ 43.90^{+  2.76}_{ -2.73}$
& $ 33.23^{+  2.40}_{ -2.43}$
& $ 24.74^{+  2.42}_{ -2.52}$
\\
${\gamma^0_X/\gamma^0_V-1}$
& $  0.62^{+  0.06}_{ -0.05}$
& $  0.33^{+  0.03}_{ -0.03}$
& $  0.00^{+  0.00}_{  0.00}$
& $ -0.24^{+  0.01}_{ -0.01}$
& $ -0.44^{+  0.03}_{ -0.03}$
\\
$\gamma^1_X$
& $ -0.29^{+  5.63}_{ -5.29}$
& $ -7.82^{+  4.70}_{ -4.52}$
& $-14.71^{+  3.79}_{ -3.78}$
& $-12.48^{+  3.21}_{ -3.23}$
& $-15.79^{+  2.98}_{ -2.99}$
\\
${\gamma^1_X/\gamma^1_V-1}$
& $ -0.98^{+  0.29}_{ -0.50}$
& $ -0.47^{+  0.15}_{ -0.25}$
& $  0.00^{+  0.00}_{  0.00}$
& $ -0.15^{+  0.04}_{ -0.04}$
& $  0.07^{+  0.18}_{ -0.12}$
\\
$\sigma_p \phi_X$
& $ 0.010^{+ 0.009}_{-0.009}$
& $-0.021^{+ 0.007}_{-0.007}$
& $-0.021^{+ 0.006}_{-0.007}$
& $-0.008^{+ 0.006}_{-0.006}$
& $ 0.036^{+ 0.007}_{-0.007}$
\\
\hline
\end{tabular}
\caption{68\% credible intervals for the global fit parameters of the Extrinsic--Intrinsic Model~II in \S\ref{modelII:sec}.\label{global2:tab}}
\end{table}

Analysis of this model produces $\gamma$ vectors that can be
transformed to the Fitzpatrick dust-extinction model with a corresponding $2.34 <R< 2.73$ 68\% credible interval.

The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  It is inconsistent with zero.  
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero, The median values
for each element are $(0.010, -0.021, -0.021, -0.008, 0.036)$.  The direction of $\phi$ is not aligned with the plane defined by
the $\gamma$ vectors; the 68\% credible interval for the  quadratic fraction of
$\phi$ in the $\gamma_0$--$\gamma_1$ plane is $0.088^{+0.118}_{-0.063}$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3/sigev.pdf} 
            \caption{Model~II posterior  contours for $\sigma_p \vec{\phi}$.           \label{M1ev:fig}}
\end{figure}

For supernovae with host-properties from Childress, we plot $p$ as a function of host-galaxy mass.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{../output_fix3/childress.pdf} 
   \caption{Intrinsic parameter
 $p$ as a function of host-galaxy mass
    \label{childress:fig}}
\end{figure}

\section{Model I and Model II Comparison}
Model~II introduces a degree of freedom that is associated with a Normal distribution.
It is thus interesting to compare the distributions of the $k_0$ and $k_1$ parameters of Model~I and Model~II,
which are shown in Figure~\ref{kcomp:fig}.  The apparent difference between Model~I and II is in the $\gamma^1k_1$ term,
where it seems that a Normal distribution component in Model~I is no longer in Model~II but is  assigned to the $\sigma_P \vec{\phi}$.
This results in a change in the direction of the fit $\gamma^1$ and shifts the inferred Fitzpatrick $R$ to higher values.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.6in]{../output_fix1/deltagamma0_med.pdf} 
   \includegraphics[width=2.6in]{../output_fix1/deltagamma1_med.pdf} 
   \includegraphics[width=2.6in]{../output_fix3/deltagamma0_med.pdf} 
   \includegraphics[width=2.6in]{../output_fix3/deltagamma1_med.pdf} 
   \caption{
   Left:
   Normalized stack of the posteriors of all supernovae  and histogram 
      of  per-supernova 
median values of $\gamma^0_{\hat{V}} k_0-\gamma^0_{\hat{V}} k_0|_0$.  Right:  The same plots for $\gamma^1_{\hat{V}} k_1-\gamma^1_{\hat{V}} k_1|_0$,
Top: Model~I. Bottom: Model~II.
    \label{kcomp:fig}}
\end{figure}

XXXXXX

\section{Supernova Model I: Two Extrinsic Parameters}
\label{model:sec}

@@

Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig} -- \ref{global5:fig}.  In these tests there is no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
For four selected colors, a plot of the
model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
 versus observed colors and the histogram of their differences
are shown in Figure~\ref{residual:fig}. 
There is no evidence of catastrophically poor model predictions nor of extreme
outlying data.
We have rerun the analysis with a variety of initial conditions, including one with $\vec{\gamma}$ equal to zero, except for a small positive 
$\gamma^0_{\hat{U}}$ and small negative $\gamma^1_{\hat{U}}$;
the 68\% credible intervals
remain consistent.  Note that the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $k$.

@@

\subsection{Per-Supernova Parameter -- Intrinsic Covariance Matrix Degeneracy}
\label{degenerate:sec}
\color{blue}
As discussed in \S\ref{modelmodel:sec}, our two models have degeneracy between per-supernova parameters and the covariance matrix.
In the posterior, the degeneracy is broken by the determinant of the covariance matrix in the likelihood and by the priors.
The posterior is thus sensitive to the prior, which being most likely biased at some level, means that the
intrinsic color parameters, $\Delta$, $k$, and $D$, and intrinsic dispersion $C_c$ are susceptible to the misassignment between each other.
The total variance and the allocation of that variance between the eigenvectors of $C_c$ in particular influence the location of the posterior
maximum, and their distributions are shown in 
\S\ref{priorappendix:sec}.  Given our particular interest in the new color term, in this Subsection we consider the possibility that
$\delta$ is an artifact of leakage from intrinsic covariance.

In our model, some intrinsic dispersion could be attributed to  $\delta$.  If this were the case, one would expect for the dominant eigenvector
of $C_c$ to be aligned with $\delta$, as the corresponding eigenvalue decreases in exchange for the
artificial creation of $D$.  No such alignment is seen: the median dot products between unit $\delta$ and the ordered five $C_c$ eigenvectors
are 0.48, 0.53,
0.34,
0.24,
0.31.
